---
title: "STAT 630 Project"
author: "Nikhil Tarigonda"
date: "2022-12-05"
output: pdf_document
---

Loading all the required libraries,

```{r, message=FALSE}
library(lubridate)
library(tidyverse)
library(readxl)
library(corrplot)
library(caTools)
library(caret)
library(wordcloud)
library(performance)
```

Loading the dataset and checking the dimensions,

```{r}
airline <- read_excel(here::here("data/Data_Train.xlsx"))
#dim(airline)
#head(airline)
#str(airline)
```
Data Cleaning:

adding duration_minutes variable from Duration variable,

```{r}
airline <- airline %>% 
  mutate(hour_min =parse_number(str_extract(airline$Duration, "[0-9]+h"))*60) %>% 
  mutate(minutes = parse_number(str_extract(airline$Duration, "[0-9]+m"))) %>% 
  mutate(minutes = replace(minutes,is.na(minutes),0)) %>% 
  mutate(duration_minutes = hour_min + minutes)
```

check if there are any na's, there are two observations which will be removed
uncomment and the observations can be removed

```{r}
which(is.na(airline$Total_Stops))
#any(is.na(airline$hour_min))
#airline <- airline[-6475,]
#airline <- airline[-9039,]
```
Histogram to check the distribution of our response(Price),

```{r}
airline %>% 
  ggplot(aes(x = Price)) + 
  geom_histogram(fill = "blue", bins= 10,col=I("black"), binwidth = 5000) + 
  theme_classic() + 
  ggtitle("Distribution of Price")
```

Using log transformation since the data is skewed,

```{r}
airline %>% 
  ggplot(aes(x = log(Price))) +
  geom_histogram(fill = "blue", bins = 10,col=I("black")) +
  theme_classic() + 
  ggtitle("Log transformed distribution of Price")
```

simple linear model,

```{r}
#model_slr <- lm(Price ~ duration_minutes, data = airline)
model_slr <- lm(log(Price) ~ duration_minutes, data = airline)
summary(model_slr)
```

Scatter plot of Price vs duration_minutes,

```{r}
airline %>% 
  ggplot(aes(x = duration_minutes, y = log(Price))) + 
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("Duration (minutes)") +
  ylab("log(Price)") +
  ggtitle("Price vs Duration") +
  theme_classic() +
  theme(plot.title = element_text(size = 30))
```

Assumptions,

```{r}
check_model(model_slr, check = c("qq","linearity","homogeneity"))
```

Adding stops as a factor,

```{r}
unique(airline$stops)
airline <- airline %>% 
  mutate(stops = case_when(
    Total_Stops == "non-stop" ~ 0,
    Total_Stops == "1 stop" ~ 1,
    Total_Stops == "2 stops" ~ 2,
    Total_Stops == "3 stops" ~ 3,
    Total_Stops == "4 stops" ~ 4))
```


multiple linear model,

```{r}
model_mlr <- lm(log(Price) ~ duration_minutes +factor(stops), data = airline)
summary(model_mlr)
```

Confidence Intervals,

```{r}
confint(model_slr)
confint(model_mlr)
```

Correlation,

```{r}
numeric_airline <- sapply(airline, is.numeric)
cor.data <- cor(airline[, numeric_airline])
corrplot.mixed(cor.data[c(1,4,5),c(1,4,5)], order = 'AOE', upper = "pie")
```

Cross validation,

```{r}
set.seed(102)
sample_airline <- sample.split(airline$Price, SplitRatio = 0.7)
# train data
train <- subset(airline, sample_airline == TRUE)
# test data
test <- subset(airline, sample_airline == FALSE)
```

Prediction,

```{r}
model_slr <- lm(log(Price) ~ duration_minutes, data = train)
model_mlr <- lm(log(Price) ~ duration_minutes +factor(stops), data = train)

price.prediction.slr <- predict(model_slr, test)
price.prediction.mlr <- predict(model_mlr, test)

```

Results,

```{r}
results <- cbind(test$Price, exp(price.prediction.slr), exp(price.prediction.mlr))
colnames(results) <- c('actual','slr_predicted','mlr_predicted')
results <- as.data.frame(results)
head(results,10)

set.seed(123)
results %>% sample_n(10)
```

